VOI-aware Monte Carlo Sampling in Trees

### 2011-12-04 paper/vago/paper.tex
paper in progress

@@

 The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) can be bounded from above using concentration
inequalities. In particular, Lemma~\ref{lemma:hoeffding-prob-bounds} is
based on the Hoeffding inequality (\ref{eqn:conc-hoeffding}):
\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) are bounded from above as
\begin{eqnarray}
\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)& \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
\Pr(X_i' \le \overline X_\beta|i=\alpha)& \le & 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
\Pr(X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
\label{eqn:probound-perf-hoeffding}
\end{eqnarray}
\label{lemma:hoeffding-prob-bounds}
\end{lmm}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{vct-wins.pdf}
\caption{Winning rate: UCT against VCT}
\label{fig:uct-against-vct}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{ect-wins.pdf}
\caption{Winning rate: UCT against ECT}
\label{fig:uct-against-ect}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bct-wins.pdf}
\caption{Winning rate: UCT against BCT}
\label{fig:uct-against-bct}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-colorful.pdf}
\caption{Best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}

>>>>

The intrinsic VOI of pulling an arm is the expected decrease
in the regret compared to selecting an arm without pulling any arm at
all. The intrinsic VOI can be estimated as the intrinsic 
value of perfect information $\Lambda_i$ about the mean reward of the $i$th arm. Two
cases are possible:
\begin{itemize}
\item the arm ($\alpha$) with the highest sample mean is pulled, and the 
mean of the arm is lower than the sample mean the second-best arm ($\beta$);
\item another arm is pulled, and the mean of the arm is higher
than the current highest sample mean ($\alpha$).
\end{itemize}
$\Lambda_i$ can be bounded from above as the probability that a different arm is selected times the
maximum possible increase in the reward:

@@

 The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) can be bounded from above using concentration
inequalities. In particular, Lemma~\ref{lemma:hoeffding-prob-bounds} is
based on the Hoeffding inequality (\ref{eqn:conc-hoeffding}):
\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) are bounded from above as
\begin{eqnarray}
\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)& \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) \le 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
\label{eqn:probound-perf-hoeffding}
\end{eqnarray}
\label{lemma:hoeffding-prob-bounds}
\end{lmm}

@@

\subsection{Selecting best arm}

Figure~\ref{fig:random-instances}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{flat.pdf}
\caption{Random instances: regret vs. number of samples}
\label{fig:random-instances}
\end{figure}

\subsection{Playing Go against UCT}

Figure~\ref{fig:uct-against-vct}, Figure~\ref{fig:uct-against-ect}, Figure~\ref{fig:uct-against-bct},
Figure~\ref{fig:best-winning-rate}.

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{vct-wins.pdf}
\caption{Go: winning rate --- UCT against VCT}
\label{fig:uct-against-vct}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{ect-wins.pdf}
\caption{Go: winning rate --- UCT against ECT}
\label{fig:uct-against-ect}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bct-wins.pdf}
\caption{Go: winning rate --- UCT against BCT}
\label{fig:uct-against-bct}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{bests-colorful.pdf}
\caption{Go: best winning rate comparison}
\label{fig:best-winning-rate}
\end{figure}
